---
title: "Australian Federal Election 2016 - Polling Place Breakdown [TENTATIVE]"
output: 
    flexdashboard::flex_dashboard:
        orientation: columns
        vertical_layout: fill
        storyboard: false
        theme: united
        social: [ "twitter" ]
        source_code: embed
        highlight: pygments
runtime: shiny
---

```{r setup, include=FALSE}
## BASED ON BREXIT VOTE FLEXDASHBOARD:
## https://github.com/p0bs/p0bs.github.io/blob/master/BrexitDashboard.Rmd
## by Robin Penfold
library(shiny)
library(gpclib)
library(rgeos)
library(flexdashboard)
library(htmltools)
library(crosstalk)
library(rmarkdown)        # Generates this html report
library(knitr)            # Underpins this html report
library(tidyr)            # Restructures data
library(magrittr)         # Enables chainings of commands
library(dplyr)            # Provides database functionality
library(htmlwidgets)      # Enables responsive exhibits
library(leaflet)          # Generates interactive map plots
library(ggplot2)
library(sp)
library(plotly)

# AEC <- rgdal::readOGR(dsn = ".", layer = "COM_ELB")
# saveRDS(AEC, file = "AEC.rds")
AEC <- readRDS(file = "AEC.rds")

# polling <- read.csv("http://vtr.aec.gov.au/Downloads/GeneralPollingPlacesDownload-20499.csv",
#                     stringsAsFactors = FALSE, skip = 1) %>% filter(!is.na(Latitude), !is.na(Longitude))
# saveRDS(polling, "polling.rds")
polling <- readRDS("polling.rds")

# results_base <- "http://vtr.aec.gov.au/Downloads/HouseStateFirstPrefsByPollingPlaceDownload-20499-"
# results.ACT <- read.csv(paste0(results_base,"ACT.csv"), stringsAsFactors = FALSE, skip = 1)
# results.NSW <- read.csv(paste0(results_base,"NSW.csv"), stringsAsFactors = FALSE, skip = 1)
# results.NT  <- read.csv(paste0(results_base,"NT.csv"),  stringsAsFactors = FALSE, skip = 1)
# results.QLD <- read.csv(paste0(results_base,"QLD.csv"), stringsAsFactors = FALSE, skip = 1)
# results.SA  <- read.csv(paste0(results_base,"SA.csv"),  stringsAsFactors = FALSE, skip = 1)
# results.TAS <- read.csv(paste0(results_base,"TAS.csv"), stringsAsFactors = FALSE, skip = 1)
# results.VIC <- read.csv(paste0(results_base,"VIC.csv"), stringsAsFactors = FALSE, skip = 1)
# results.WA  <- read.csv(paste0(results_base,"WA.csv"),  stringsAsFactors = FALSE, skip = 1)
# results <- bind_rows(mget(ls(pattern = "results\\.+.*")))
# saveRDS(results, "results.rds")
results <- readRDS("results.rds")

# election <- merge(results, polling)
# saveRDS(election, "election.rds")
election <- readRDS("election.rds")

election %<>% group_by(PollingPlaceID) %>% 
    arrange(-OrdinaryVotes) %>% 
    mutate(PPwinner    = first(PartyAb), 
           PPvotes     = sum(OrdinaryVotes), 
           Ncandidates = n_distinct(CandidateID)) %>% 
    filter(PPvotes > 0)
election %<>% 
    group_by(DivisionID) %>% 
    arrange(-OrdinaryVotes) %>% 
    mutate(ELECwinner = first(PartyAb), 
           ELECvotes  = sum(OrdinaryVotes))

election.CT <- SharedData$new(election)

# election2PP <- read.csv("http://vtr.aec.gov.au/Downloads/HouseTppByPollingPlaceDownload-20499.csv", 
                        # stringsAsFactors = FALSE, skip = 1)
# election2PPall <- merge(election2PP, polling)
# saveRDS(election2PPall, file="election2PPall.rds")
election2PP <- readRDS("election2PPall.rds")

election2PP %<>% rename(LP = Liberal.National.Coalition.Votes,
                        ALP = Australian.Labor.Party.Votes)
election2PP %<>% 
    group_by(PollingPlaceID) %>% 
    mutate(PPwinner = ifelse(LP + ALP == 0, "NA", 
                             ifelse(LP > ALP, "LP", 
                                    ifelse(LP == ALP, "EVEN", "ALP"))))
election2PP %<>% 
    group_by(DivisionID) %>% 
    summarise(ElecLPVotes = sum(LP), 
              ElecALPVotes = sum(ALP),
              ELECwinner = ifelse(ElecLPVotes + ElecALPVotes == 0, "NA", 
                                  ifelse(ElecLPVotes > ElecALPVotes, "LP", "ALP"))) %>% 
    merge(election2PP)

election2PP.CT <- SharedData$new(election2PP)

partycols <- c("blue","red","darkgreen","orange","green","darkblue",
               "skyblue","brown","black","yellow","darkred","grey",
               "purple","white","grey","grey","grey")
partylabs <- c("LP","ALP","NP","XEN","GRN","LNP","CLP","KAP","ON",
               "DHJP","ARF","AJP","IND","INF","CDP","FFP","LDP")

dfpp <- reactive({
    
    df_raw <- election.CT$data(withSelection = TRUE) %>% filter(StateAb %in% input$state)
    
    df <- df_raw %>% 
        group_by(DivisionNm, PollingPlaceNm) %>% 
        summarise(PPwinner = unique(PPwinner),
                  nVotes   = sum(OrdinaryVotes),
                  Latitude = unique(Latitude),
                  Longitude = unique(Longitude))
    
    df$PPwinner <- factor(df$PPwinner, levels = partylabs)
    df <- arrange(df, nVotes)
    
    df
    
})

dfpp2PP <- reactive({
    
    df_raw <- election2PP.CT$data(withSelection = TRUE) %>% filter(StateAb %in% input$state2PP)
    
    df <- df_raw
    
    df$PPwinner <- factor(df$PPwinner, levels = c("LP","ALP","EVEN","NA"))
    
    df
    
})

dfrawvotes <- reactive({
    
    df_raw <- election.CT$data(withSelection = TRUE) %>% filter(StateAb %in% input$state)
    
    df <- df_raw %>% 
        group_by(DivisionNm, PartyAb) %>% 
        summarise(nVotes   = sum(OrdinaryVotes))
    
    df$PartyAb <- factor(df$PartyAb, levels = partylabs)
    df <- arrange(df, PartyAb)
    
    df    
    
})

dfrawvotes2PP <- reactive({
    
    df_raw <- election2PP.CT$data(withSelection = TRUE) %>% filter(StateAb %in% input$state2PP)
    
    df <- df_raw %>% gather(Party, Votes, LP, ALP)
    
    df$Party <- factor(df$Party, levels = c("LP", "ALP","EVEN","NA"))
    
    df    
    
})


dfswing <- reactive({
    
    df_raw <- election.CT$data(withSelection = TRUE) %>% filter(StateAb %in% input$state)
    
    df <- df_raw %>%
        group_by(DivisionNm, PollingPlaceID, PartyAb) %>%
        summarise(swing     = mean(Swing),
                  PPwinner  = unique(PPwinner))
    
    df$PPwinner <- factor(df$PPwinner, levels = partylabs)
    
    df
    
})

dfAEC <- reactive({
    
    newAEC <- AEC[AEC$State == input$state, ]
    newAEC@data <- merge(AEC@data %>% filter(State %in% input$state), 
                         election %>% filter(StateAb %in% input$state) %>% 
                             group_by(DivisionNm) %>% 
                             summarise(ELECwinner = unique(ELECwinner)), 
                         by.x = "Elect_div", by.y = "DivisionNm")
    
    newAEC
    
})

dfAEC2PP <- reactive({
    
    newAEC <- AEC[AEC$State == input$state2PP, ]
    newAEC@data <- merge(AEC@data %>% filter(State %in% input$state2PP), 
                         election2PP %>% filter(StateAb %in% input$state2PP) %>% 
                             group_by(DivisionNm) %>% 
                             summarise(ELECwinner = unique(ELECwinner)), 
                         by.x = "Elect_div", by.y = "DivisionNm")
    
    newAEC
    
})


```

First Preferences
==================

Column {.sidebar}
-----------------------------------------------------------------------

### 

```{r}
selectInput("state", label = "Select which State/Territory to visualise", choices = levels(AEC$State), selected = "SA")
```

© Commonwealth of Australia 2016

Election results obtained from the [AEC Tally Room](http://vtr.aec.gov.au/HouseDownloadsMenu-20499-Csv.htm). The Australian Electoral Commission ([AEC](http://www.aec.gov.au)) is an independent statutory authority that is politically neutral and does not sponsor, endorse or approve of material found on sites that link to the AEC website (www.aec.gov.au).

[Electoral spatial boundaries](http://www.aec.gov.au/Electorates/gis/gis_datadownload.htm) obtained from the Australian Electoral Commission ([AEC](http://www.aec.gov.au)) and are © Commonwealth of Australia (Australian Electoral Commission) 2016.

<br />

Created with [flexdashboard](https://github.com/rstudio/flexdashboard/) by [Jonathan Carroll](http://jcarroll.com.au). A copy of the source for this page can be found [here](http://github.com/jonocarroll/) (feedback welcome). [Share this!](https://twitter.com/intent/tweet?text=Australian%20Federal%20Election%202016%20-%20Polling%20Place%20Breakdown%20using%20flexdashboard&url=http%3A%2F%2Fjcarroll.shinyapps.io%2FAusElection2016&original_referer=http%3A%2F%2Fjcarroll.shinyapps.io%2FAusElection2016)

```{r}
renderImage({list(
    src = "cc-icon.png",
    contentType = "image/png",
    alt = "CC-BY-3.0")}, 
    deleteFile = FALSE
)
```

Column {.tabset .tabset-fade} 
-----------------------------------------------------------------------

### Swing

```{r fig.height=5, fig.width=5}

renderPlot({
    
    myCols <- partycols
    names(myCols) <- levels(dfpp()$PPwinner)
    
    gg <- ggplot(dfswing(), aes(x = DivisionNm, y = swing)) + 
        geom_point(aes(col = PPwinner), alpha = 0.5, position = "jitter") +
        scale_color_manual(values = myCols, name = "PP Winner") +
        theme_minimal() +
        theme(axis.text.x = element_text(angle = 90)) +
        labs(x = "Division", y = "Swing at each Polling Place")
    
    print(gg)
    
})

```

### Votes per Polling Place 

```{r, fig.height=5, fig.width=5}

renderPlotly({
    plot_ly(dfpp(), 
            type = "scatter", 
            x = PollingPlaceNm, 
            y = nVotes, 
            text = paste("(",DivisionNm,") ",nVotes," Votes"),
            mode = "markers") %>%
        layout(dragmode = "select", 
               xaxis = list(title = "Polling Place", showticklabels = FALSE), 
               yaxis = list(tickformat = '0.0f'))
})

```

### Votes Tallied by Polling Places

```{r fig.height=8, fig.width=5}

renderPlot({
    
    myCols <- partycols    
    names(myCols) <- levels(dfpp()$PPwinner)
    
    gg <- ggplot(dfpp(), aes(x = DivisionNm)) +
        scale_fill_manual(values = myCols, name = "PP Winner") +
        scale_alpha_manual(values = c(1,0.25), guide = FALSE) +
        geom_bar(stat = "count", aes(fill = PPwinner)) + 
        theme_minimal() + 
        labs(y = "Number of Polling Places Won by Party") + 
        theme(axis.text.x = element_text(angle = 90))
    
    print(gg)
    
    
})

```

### Raw Votes

```{r fig.height=8, fig.width=5}

renderPlot({
    
    myCols <- partycols    
    names(myCols) <- levels(dfrawvotes()$PartyAb)
    
    gg <- ggplot(dfrawvotes(), aes(x = DivisionNm)) +
        scale_fill_manual(values = myCols, name = "Party") +
        scale_alpha_manual(values = c(1,0.25), guide = FALSE) +
        geom_bar(stat = "identity", position = "fill", aes(y = nVotes, fill = PartyAb)) +
        theme_minimal() + 
        labs(y = "Raw Votes Won by Party") + 
        theme(axis.text.x = element_text(angle = 90))
    
    print(gg)
    
})

```

Column
-----------------------------------------------------------------------

### Mapped Polling Places

```{r fig.height=7, fig.width=8}

renderLeaflet({
    
    rAEC  <- dfAEC()
    rdfpp <- dfpp()
    
    factpal  <- colorFactor(partycols , factor(rAEC@data$ELECwinner, levels = partylabs))
    factpal2 <- colorFactor(partycols , factor(rdfpp$PPwinner, levels = partylabs))

    # rdfpp$PPvotesF <- cut(rdfpp$nVotes, breaks = c(0,100,500,1000,2000,5000,10000,20000), right = FALSE, labels = c(1,6,8,10,12,14,16) )

    map <- leaflet(data = rdfpp) %>% 
        addProviderTiles("CartoDB.Positron", 
                         options = providerTileOptions(opacity = 0.6)) %>%
        addPolygons(data = rAEC, 
                    fillColor = ~factpal(ELECwinner), 
                    color = "grey", 
                    weight = 3, 
                    popup = ~paste0("Electorate: ",Elect_div)) %>% 
        setView(lng = median(rdfpp$Longitude), 
                lat = median(rdfpp$Latitude), 
                zoom = 11) %>%
        addCircleMarkers(lat = ~Latitude, 
                         lng = ~Longitude, 
                         popup = ~PollingPlaceNm, 
                         color = ~factpal2(PPwinner), 
                         radius = 6,
                         # radius = ~PPvotesF, 
                         stroke = FALSE, 
                         fillOpacity = 0.9) %>% 
        addLegend(colors = partycols, 
                  labels = partylabs) 
    
    map
    
})

```

Two Party Preferred
==================

Column {.sidebar}
-----------------------------------------------------------------------

### 

```{r}
selectInput("state2PP", label = "Select which State/Territory to visualise", choices = levels(AEC$State), selected = "SA")
```

© Commonwealth of Australia 2016

Election results obtained from the [AEC Tally Room](http://vtr.aec.gov.au/HouseDownloadsMenu-20499-Csv.htm). The Australian Electoral Commission ([AEC](http://www.aec.gov.au)) is an independent statutory authority that is politically neutral and does not sponsor, endorse or approve of material found on sites that link to the AEC website (www.aec.gov.au).

[Electoral spatial boundaries](http://www.aec.gov.au/Electorates/gis/gis_datadownload.htm) obtained from the Australian Electoral Commission ([AEC](http://www.aec.gov.au)) and are © Commonwealth of Australia (Australian Electoral Commission) 2016.

<br />

Created with [flexdashboard](https://github.com/rstudio/flexdashboard/) by [Jonathan Carroll](http://jcarroll.com.au). A copy of the source for this page can be found [here](http://github.com/jonocarroll/) (feedback welcome). [Share this!](https://twitter.com/intent/tweet?text=Australian%20Federal%20Election%202016%20-%20Polling%20Place%20Breakdown%20using%20flexdashboard&url=http%3A%2F%2Fjcarroll.shinyapps.io%2FAusElection2016&original_referer=http%3A%2F%2Fjcarroll.shinyapps.io%2FAusElection2016)

```{r}
renderImage({list(
    src = "cc-icon.png",
    contentType = "image/png",
    alt = "CC-BY-3.0")}, 
    deleteFile = FALSE
)
```

Column {.tabset .tabset-fade} 
-----------------------------------------------------------------------

### Votes Tallied by Polling Places

```{r fig.height=8, fig.width=5}

renderPlot({
    
    myCols <- c("blue","red","grey","#D3D3D3")   
    names(myCols) <- c("LP","ALP","EVEN","NA")
    
    gg <- ggplot(dfpp2PP(), aes(x = DivisionNm)) +
        scale_fill_manual(values = myCols, name = "PP Winner") +
        scale_alpha_manual(values = c(1,0.25), guide = FALSE) +
        geom_bar(stat = "count", aes(fill = PPwinner, alpha = selected_)) + 
        theme_minimal() + 
        labs(y = "Number of Polling Places Won by Party") + 
        theme(axis.text.x = element_text(angle = 90))
    
    print(gg)
    
})

```

### Raw Votes

```{r fig.height=8, fig.width=5}

renderPlot({
    
    myCols <- c("blue","red","#D3D3D3")   
    names(myCols) <- c("LP","ALP","NA")
    
    gg <- ggplot(dfrawvotes2PP(), aes(x = DivisionNm)) +
        scale_alpha_manual(values = c(1,0.25), guide = FALSE) +
        geom_bar(stat = "identity", position = "fill", aes(y = Votes, fill = Party)) +
        scale_fill_manual(values = myCols, name = "Party") +
        theme_minimal() + 
        labs(y = "Raw Votes Won by Party") + 
        theme(axis.text.x = element_text(angle = 90))
    
    print(gg)
    
})

```

Column
-----------------------------------------------------------------------

### Mapped Polling Places

```{r fig.height=7, fig.width=8}

renderLeaflet({
    
    rAEC  <- dfAEC2PP()
    rdfpp <- dfpp2PP()
    
    factpal  <- colorFactor(c("blue","red","grey","#D3D3D3"), factor(rAEC@data$ELECwinner, levels = c("LP","ALP","EVEN","NA")))
    factpal2 <- colorFactor(c("blue","red","grey","#D3D3D3"), factor(rdfpp$PPwinner, levels = c("LP","ALP","EVEN","NA")))
    
    map <- leaflet(data = rdfpp) %>% 
        addProviderTiles("CartoDB.Positron", 
                         options = providerTileOptions(opacity = 0.6)) %>%
        addPolygons(data = rAEC, 
                    fillColor = ~factpal(ELECwinner), 
                    color = "grey", 
                    weight = 3, 
                    popup = ~paste0("Electorate: ",Elect_div)) %>% 
        setView(lng = median(rdfpp$Longitude), 
                lat = median(rdfpp$Latitude), 
                zoom = 11) %>%
        addCircleMarkers(lat = ~Latitude, 
                         lng = ~Longitude, 
                         popup = ~PollingPlaceNm, 
                         color = ~factpal2(PPwinner), 
                         radius = 6, 
                         stroke = FALSE, 
                         fillOpacity = 0.9) %>% 
        addLegend(colors = c("blue","red","grey","#D3D3D3"), 
                  labels = c("LP","ALP","EVEN","NA"))
    
    map
    
})

```

